### 2021-02-11 TCGA_UCEC_Data_Analysis.Rmd

---
title: "Assessment of Normalization Methods for TCGA-UCEC Data"
author: "Yannick Dueren, Johannes Lederer, Li-Xuan Qin"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
# abstract: "TODO: This is the abstract."
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "results") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

The settings list `settings` configure which parts, simulations, and analysis steps of this R markdown are executed. This way single components of the analysis can be disabled for the purpose of saving computation time. Setting `settings$all` to `TRUE` will set all other options to true regardless of their value.

```{r analysis.settings}
settings <- list(
  # If true sets all other settings to TRUE
  all=F,
  # Application of the full DANA pipeline to the benchmark data
  bench.DANA=T,
  # Application of the full DANA pipeline to the test data
  test.DANA=T,
  # Generate interactive plots of co-expression structures
  investigate.coexpression=F,
  # Perform Differential Expression Analysis for test and benchmark data sets
  perform.DEA=T,
  # Generate paper figures
  generate.paper.figs=T,
  # Specify if paper figures are exported
  export.figures = T,
  # Path for file exports
  paper.fig.path = "../../danawriteup/figs/",
  # Clears environment
  debug=TRUE
)
if(settings$all) {
  settings$bench.DANA=TRUE
  settings$test.DANA=TRUE
  settings$investigate.coexpression=TRUE
  settings$generate.paper.figs=TRUE
}
if(settings$generate.paper.figs) {
  settings$bench.DANA=TRUE
  settings$test.DANA=TRUE
}
if(settings$debug) rm(list=ls()[ls()!="settings"])
```

## Parameter configuration

The `config` list contains all parameters for the analysis.
Remember to *always set the working directory* and the paths to the data/external files prior to running the code.

```{r configuration}
config <- list(
  # working directory
  working.dir = "C:/Users/Yannick/Google Drive/projects/normalization/paperSupplement",
  DANA.path = "./R/DANA.R",

  ## Data
  case.name = "UCEC",
  # Read count data file for the full UCEC data set
  data.file.path = "data/TCGA_harmonized_UCEC.RData",
  # RData file for coordinates data frame based on miRBase miRNA definitions
  coords.file.path = "data/miRBase.coords.RData",

  ## Normalization Methods
  # Specify which normalization methods will be applied
  norm.apply.TC = TRUE,
  norm.apply.UQ = TRUE,
  norm.apply.med = TRUE,
  norm.apply.TMM = TRUE,
  norm.apply.DESeq = TRUE,
  norm.apply.PoissonSeq = TRUE,
  norm.apply.QN = TRUE,
  norm.apply.RUV = TRUE,

  # thresholds for zero-expressed, poorly-expressed and well-expressed genes 
  t.zero = 2,  # zero-expressed in [0, t.zero)
  t.poor = 5,  # poorly-expressed in [t.zero, t.well]
  t.well = 2^6,  # well-expressed in [t.well, inf)
  # distance threshold for clustering
  cluster.threshold = 10000,
  # preprocessing data transformation type: none ("n"), modified log2 ("log2"),
  # or cube root ("cube") available
  preprocess.transform = "log2",

  ## Correlation Estimation for positive and negative controls
  # Available methods | Tuning parameter calibration schemes
  # "mb"              | "cv", "aic", "bic", "av"
  # "huge.mb"         | "ric", "stars"
  # "glasso"          | "ric", "stars", "ebic"
  # "fastggm"         | none
  # "pearson"         | none
  # "spearman"        | none

  # Positive Controls
  corr.method.pos = "mb",
  tuning.criterion.pos = "bic",
  # Negative Controls
  corr.method.neg = "pearson",
  tuning.criterion.neg = "",
  # Generate plots within DANA
  generate.plots = FALSE  # generate comparison plots
)
```

Configuration for the differential expression analysis

```{r configuration.DEA}
config.DEA <- list(
  ## Data
  case.name = "UCEC",

  ## Normalization Methods
  # Specify which normalization methods will be applied
  norm.apply.TC = TRUE,
  norm.apply.UQ = TRUE,
  norm.apply.med = TRUE,
  norm.apply.TMM = TRUE,
  norm.apply.DESeq = TRUE,
  norm.apply.PoissonSeq = TRUE,
  norm.apply.QN = TRUE,
  norm.apply.RUV = TRUE,

  # Significance level for DEA
  alpha = 0.01,
  # Plots
  generate.volcano.plot = TRUE,
  generate.meanvar.plot = TRUE,
  # Use q-values (adjusted p-values) instead of p-values
  q.values = FALSE,
  # RUV reduces the parameter size. Reduce DEA result to RUV genes?
  perform.subsetting = FALSE
)
```

## Load Packages

Load all required R libraries/packages.

```{r load.packages, results='hide', message=FALSE}
# Set working directory
setwd(config$working.dir)
# DANA functions
source(config$DANA.path)
# Libraries
library(openxlsx)  # read excel files
library(corrplot) # For mixed correlation plots
library(cowplot) # Arrange plots
library(RColorBrewer) # Colors
library(latex2exp) # for latex in axis labels 
library(ggcorrplot) # ggplot2 correlation plots
```

# Load Data

## Load the dataset under study

We load the data set into memory.

```{r load.data.set}
# Load the p x n matrix of read counts into the workspace
load(config$data.file.path)

# Unlist the data
benchmark <- TCGA.UCEC$benchmark
test <- TCGA.UCEC$test
benchmark.end <- benchmark$END
benchmark.ser <- benchmark$SER
test.end <- test$END
test.ser <- test$SER

# Transform gene names to lower case
rownames(benchmark.end) <- tolower(rownames(benchmark.end))
rownames(benchmark.ser) <- tolower(rownames(benchmark.ser))
rownames(test.end) <- tolower(rownames(test.end))
rownames(test.ser) <- tolower(rownames(test.ser))

# Rename
test.RC <- test <- cbind(test.end, test.ser) 
test.groups <- c(rep("END", dim(test.end)[2]), rep("SER", dim(test.ser)[2]))
bench.RC <- benchmark <- cbind(benchmark.end, benchmark.ser)  
bench.groups <- c(rep("END", dim(benchmark.end)[2]), rep("SER", dim(benchmark.ser)[2]))

# Inspect
cat("Dimensions of test (mixed-batch) data: ", dim(test.RC), "\n")
cat("Dimensions of benchmark (single-batch) data: ", dim(bench.RC), "\n")
```

## miRNA Coordinates

### Load corresponding miRNA chromosome coordinates

A coordinates data frame that specifies the base pair location of each miRNA of the data set on the chromosomes is loaded.

```{r load.tcga.coords.data.frame}
# Load miRNA chromosome location coordinates data frame
load(config$coords.file.path)
coords <- coords  # change according to the name of the loaded data matrix
```

### Remove genes that cannot be found in the coordinates data frame

Some miRNAs map to multiple locations of sequence families. We named the sequence family with a parenthesis that reflects the number of members from all the different genomic locations (e.g. hsa-let-7a(3)). These miRNAs cannot be uniquely mapped to the genome, therefore we must exclude these from the location based analysis.

```{r remove.genes.not.in.coords}
# only consider genes that are present in the coordinate data frame
genes.in.coords <- rownames(coords)[na.omit(match(rownames(test.RC), rownames(coords)))]
cat(dim(test.RC)[1]-length(genes.in.coords), " genes not found in coords. Reducing data set to ", length(genes.in.coords), "genes.\n")
# test data set
test.RC <- test.RC[genes.in.coords, ]
# benchmark data set
bench.RC <- bench.RC[genes.in.coords, ]
genes <- rownames(test.RC)
rm(genes.in.coords)
```

## Examine the Data

First, we investigate the distribution of mean miRNA expression of the data.

```{r miRNA.expressions}
# Histogram plot test data set
par(mfrow=c(1,1))
df <- data.frame(log.expression=log2(rowMeans(test.RC)+1))
print(ggplot(df, aes(x=log.expression)) + 
        geom_histogram(binwidth = .1, color="black", fill="black") +
        geom_vline(xintercept = log2(config$t.zero+1), color="blue", linetype="dashed") +
        geom_vline(xintercept = log2(config$t.poor+1), color="blue", linetype="dashed") +
        geom_vline(xintercept = log2(config$t.well+1), color="red",  linetype="dashed") +
        ggtitle("mixed-batch data set"))
rm(df)

# Histogram plot benchmark data set
par(mfrow=c(1,1))
df <- data.frame(log.expression=log2(rowMeans(bench.RC)+1))
print(ggplot(df, aes(x=log.expression)) + 
        geom_histogram(binwidth = .1, color="black", fill="black") +
        geom_vline(xintercept = log2(config$t.zero+1), color="blue", linetype="dashed") +
        geom_vline(xintercept = log2(config$t.poor+1), color="blue", linetype="dashed") +
        geom_vline(xintercept = log2(config$t.well+1), color="red",  linetype="dashed") +
        ggtitle("single-batch data set"))
rm(df)
```

We observe that there is a large proportion of poorly-expressed genes. Some of them have extremely low or zero mean expression which corresponds to essentially zero reads.

# DANA - Performance Analysis of Normalization Methods

We apply the full analysis pipeline to the data set. This includes:

1. Apply the normalization methods to the raw counts.
2. Define positive and negative controls based on read counts.
3. Estimate the level of coexpression between miRNAs for each data set.
4. Compare positive and negative controls before and after normalization.

The following parameters are used:

* Case name: `r config$case.name`
* Control definition:
  - Definition of negative controls, read count in [`r config$t.zero`, `r config$t.poor`]
  - Definition of positive controls, read count in [`r config$t.well`, inf]
  - Distance threshold for clustering: `r config$cluster.threshold`
* Data preprocessing transformation: `r config$preprocess.transform`
* Coexpression estimation:
  - Correlation estimation method for positive controls: `r config$corr.method.pos`
  - Correlation estimation tuning parameter calibration method for positive controls: `r config$tuning.criterion.pos`
* Comparison statistics:
  - Generate plots? `r config$generate.plots`

## Benchmark (single-batch) Data Set

```{r bench.full.pipeline}
if(settings$bench.DANA) {
  genes <- rownames(bench.RC)

  ## Apply Normalization
  data.norm <- normalize(X=bench.RC,
                         groups=bench.groups,
                         name="benchmark",
                         apply.TC=config$norm.apply.TC,
                         apply.UQ=config$norm.apply.UQ,
                         apply.med=config$norm.apply.med,
                         apply.TMM=config$norm.apply.TMM,
                         apply.DESeq=config$norm.apply.DESeq,
                         apply.PoissonSeq=config$norm.apply.PoissonSeq,
                         apply.QN=config$norm.apply.QN,
                         apply.RUV=config$norm.apply.RUV)
  bench.norm <- data.norm

  ## Define Controls
  pre.res <- define.controls(bench.RC,
                             t.zero=config$t.zero,
                             t.poor=config$t.poor,
                             t.well=config$t.well,
                             t.cluster=config$cluster.threshold,
                             coords=coords,
                             clusters=NULL)
  pos.controls <- pre.res$genes.pos
  pos.controls.bench <- pre.res$genes.pos
  neg.controls <- pre.res$genes.neg
  neg.controls.bench <- pre.res$genes.neg
  clusters <- pre.res$clusters
  clusters.bench <- clusters

  # Apply DANA pipeline
  DANA.bench <- DANA(data.RC=bench.RC, 
                     data.norm=bench.norm, 
                     pos.controls=pos.controls.bench, 
                     neg.controls=neg.controls.bench, 
                     clusters=clusters.bench,
                     coords=coords,
                     case.name="benchmark",
                     generate.plots=config$generate.plots,
                     preprocess.transform=config$preprocess.transform,
                     corr.method.pos=config$corr.method.pos,
                     tuning.criterion.pos=config$tuning.criterion.pos,
                     corr.method.neg=config$corr.method.neg,
                     tuning.criterion.neg=config$tuning.criterion.neg)

  if(!config$generate.plots) {
    stargazer(DANA.bench$metrics, type="text", summary=FALSE, digits=4, 
              title="Comparison metrics for the single-batch data set", align=TRUE)
  }
  iplot.bench <- iggplot.corr(DANA.bench$data.model$pos$corr, clusters=clusters, title="Single-batch data set - Positive controls (un-normalized)", coords=coords)
  iplot.bench.TMM <- iggplot.corr(DANA.bench$norm.models$benchmark.TMM$pos$corr, clusters=clusters, title="Single-batch data set - Positive controls (TMM)", coords=coords)
}
```

## Test (Mixed-Batch) Data Set

```{r test.full.pipeline}
if(settings$test.DANA) {
  genes <- rownames(test.RC)

  ## Apply Normalization
  data.norm <- normalize(test.RC,
                         groups=test.groups,
                         name="test",
                         apply.TC=config$norm.apply.TC,
                         apply.UQ=config$norm.apply.UQ,
                         apply.med=config$norm.apply.med,
                         apply.TMM=config$norm.apply.TMM,
                         apply.DESeq=config$norm.apply.DESeq,
                         apply.PoissonSeq=config$norm.apply.PoissonSeq,
                         apply.QN=config$norm.apply.QN,
                         apply.RUV=config$norm.apply.RUV)
  test.norm <- data.norm

  # Define Controls  
  pre.res <- define.controls(test.RC,
                             t.zero=config$t.zero,
                             t.poor=config$t.poor,
                             t.well=config$t.well,
                             t.cluster=config$cluster.threshold,
                             coords=coords,
                             clusters=NULL)
  pos.controls <- pre.res$genes.pos
  pos.controls.test <- pre.res$genes.pos
  neg.controls <- pre.res$genes.neg
  neg.controls.test <- pre.res$genes.neg
  clusters <- pre.res$clusters
  clusters.test <- clusters

  # Apply DANA pipeline
  DANA.test <- DANA(data.RC=test.RC, 
                    data.norm=test.norm, 
                    pos.controls=pos.controls.test, 
                    neg.controls=neg.controls.test, 
                    clusters=clusters.test,
                    coords=coords,
                    case.name="test",
                    generate.plots=config$generate.plots,
                    preprocess.transform=config$preprocess.transform,
                    corr.method.pos=config$corr.method.pos,
                    tuning.criterion.pos=config$tuning.criterion.pos,
                    corr.method.neg=config$corr.method.neg,
                    tuning.criterion.neg=config$tuning.criterion.neg)

  if(!config$generate.plots) {
    stargazer(DANA.test$metrics, type="text", summary=FALSE, digits=4, 
              title="Comparison metrics for the mixed-batch data set", align=TRUE)
  }
  iplot.test <- iggplot.corr(DANA.test$data.model$pos$corr, clusters=clusters, title="Mixed-batch data set - Positive controls (un-normalized)", coords=coords)
  iplot.test.TMM <- iggplot.corr(DANA.test$norm.models$test.TMM$pos$corr, clusters=clusters, title="Mixed-batch data set - Positive controls (TMM)", coords=coords)
}
```

## Investigate Co-expression Structures

We compare the estimated co-expression structures for the mixed-batch and single-batch dataset. Here, we take a look at the models for the un-normalized data as well as TMM-normalized data. Note that you cannot directly compare the graphs of benchmark with test data since the considered set of genes is not identical.

```{r investigate.coexpression}
if(settings$investigate.coexpression && settings$bench.DANA && settings$test.DANA) {
  htmltools::tagList(list(iplot.bench, iplot.bench.TMM, iplot.test, iplot.test.TMM))  # show plots in markdown
}
```

# Differential Expression Analysis

## Normalize Data

```{r DEA.normalize.data}
if(settings$perform.DEA) {
  ## Reset data
  test.RC <- test 
  bench.RC <- benchmark

  ## Normalize test Data
  test.norm <- normalize(test.RC,
                         groups=test.groups,
                         name="test",
                         apply.TC=config.DEA$norm.apply.TC,
                         apply.UQ=config.DEA$norm.apply.UQ,
                         apply.med=config.DEA$norm.apply.med,
                         apply.TMM=config.DEA$norm.apply.TMM,
                         apply.DESeq=config.DEA$norm.apply.DESeq,
                         apply.PoissonSeq=config.DEA$norm.apply.PoissonSeq,
                         apply.QN=config.DEA$norm.apply.QN,
                         apply.RUV=FALSE)

  test.norm.RUV.adjust <- list(test.RUVr=norm.RUV(test.RC, test.groups, "RUVr")$adjust.factor,
                               test.RUVs=norm.RUV(test.RC, test.groups, "RUVs")$adjust.factor,
                               test.RUVg=norm.RUV(test.RC, test.groups, "RUVg")$adjust.factor)
}
```

## Benchmark (single-batch) DEA

```{r DEA.benchmark}
if(settings$perform.DEA) {
  ## Perform DEA on benchmark dataset
  bench.DEA <- DE.voom(data.RC=bench.RC, groups=bench.groups, plot=config.DEA$generate.meanvar.plot)
  if(config.DEA$generate.volcano.plot) plot.DE.volcano(bench.DEA, alpha=config.DEA$alpha, q.values=config.DEA$q.values)
}
```

## Test (mixed-batch) DEA

```{r DEA.test}
if(settings$perform.DEA) {
  ## Perform DEA on full dataset
  test.DEA <- DE.voom(data.RC=test.RC, groups=test.groups, plot=config.DEA$generate.meanvar.plot, plot.title="test (no norm)")
  if(config.DEA$generate.volcano.plot) plot.DE.volcano(test.DEA, alpha=config.DEA$alpha, q.values=config.DEA$q.values, title="test (no norm)")

  ## DEA for normalized test data
  test.norm.DEA <- list()
  for(i in 1:length(test.norm)) {
    test.norm.DEA <- 
      append(test.norm.DEA, list(DE.voom(data.RC=test.norm[[i]], 
                                         groups=test.groups, 
                                         plot=config.DEA$generate.meanvar.plot,
                                         plot.title=names(test.norm)[i])))
    if(config.DEA$generate.volcano.plot) {
      plot.DE.volcano(test.norm.DEA[[i]], 
                      alpha=config.DEA$alpha, 
                      q.values=config.DEA$q.values, 
                      title=names(test.norm)[i])
    }
  }
  ## DEA for RUV normalization
  for(i in 1:length(test.norm.RUV.adjust)) {
    test.norm.DEA <- 
      append(test.norm.DEA, list(DE.voom(data.RC=test.RC, 
                                         groups=test.groups, 
                                         plot=config.DEA$generate.meanvar.plot,
                                         plot.title=names(test.norm.RUV.adjust)[i],
                                         adjust=test.norm.RUV.adjust[[i]])))
    if(config.DEA$generate.volcano.plot) {
      plot.DE.volcano(test.norm.DEA[[i]], 
                      alpha=config.DEA$alpha, 
                      q.values=config.DEA$q.values, 
                      title=names(test.norm.RUV.adjust)[i])
    }
  }
  names(test.norm.DEA) <- c(names(test.norm), names(test.norm.RUV.adjust))
}
```

## Results

```{r DEA.Test.vs.Benchmark}
if(settings$perform.DEA) {
  ## Compare test (non-norm) to benchmark
  test.norm.DEA.stats <- list(test=compare.DEAs(DEA=test.DEA,
                                           truth.DEA=bench.DEA,
                                           alpha=config.DEA$alpha))
  ## Compare normalized test data to benchmark
  for(i in 1:length(test.norm.DEA)) {
    test.norm.DEA.stats <- 
      append(test.norm.DEA.stats, 
             list(compare.DEAs(DEA=test.norm.DEA[[i]],
                               truth.DEA=bench.DEA,
                               alpha=config.DEA$alpha,
                               subset=switch(config.DEA$perform.subsetting+1,
                                             NULL,
                                             rownames(test.norm.DEA[[i]])))))
  }
  names(test.norm.DEA.stats) <- c("test.No Norm", names(test.norm.DEA))
  test.norm.DEA.stats <- test.norm.DEA.stats[1:length(test.norm.DEA.stats)]

  ## Visualize results

  # FNR-FDR Plot
  test.DEA.res <- data.frame(method=substr(names(test.norm.DEA.stats), 6, 20),
                             FDR=-sapply(test.norm.DEA.stats, function(x) x$FDR)+1,
                             FNR=-sapply(test.norm.DEA.stats, function(x) x$FNR)+1)
  p.test.DEA.FNR.FDR <- ggplot(test.DEA.res, aes(x=FNR, y=FDR, label=method)) +
      theme_bw() + 
      geom_point(alpha=1) + 
      xlab("Sensitivity") + # True positive rate(1-FNR)
      ylab("Precision") + # Positive predictive value (1-FDR)
      geom_text_repel(aes(label = method), size=3) +
      scale_x_continuous(labels = scales::percent, limits=c(0,1),
                         breaks = scales::pretty_breaks(n = 4)) +
      scale_y_continuous(labels = scales::percent, limits = c(0,1))
  print(p.test.DEA.FNR.FDR)

  # TPR-FPR Plot
  test.DEA.res <- data.frame(method=substr(names(test.norm.DEA.stats), 6, 20),
                             TPR=sapply(test.norm.DEA.stats, function(x) x$TPR),
                             FPR=-sapply(test.norm.DEA.stats, function(x) x$FPR)+1)
  p.test.DEA.TPR.FPR <- ggplot(test.DEA.res, aes(x=TPR, y=FPR, label=method)) + 
      theme_bw() + 
      geom_point(alpha=1) + 
      xlab("TPR") + 
      ylab("1-FPR") + 
      geom_text_repel(aes(label = method), size=3) +
      scale_x_continuous(labels = scales::percent, limits=c(0,1),
                         breaks = scales::pretty_breaks(n = 4)) +
      scale_y_continuous(labels = scales::percent, limits = c(0,1))
  print(p.test.DEA.TPR.FPR)

  # CAT Plot
  d <- append(list("test.No Norm"=test.DEA), test.norm.DEA)
  names(d) <- substr(names(d), 6, 20)
  p.test.DEA.CAT <- plot.CAT(DEA=d, 
                             truth.DEA=bench.DEA, 
                             title="Significance ranks of miRNAs",
                             maxrank=50)
  print(p.test.DEA.CAT)
  rm(d)
}
```

# Results and Figures for Paper

## Number of samples, markers, and positive and negative controls

```{r generate.paper.figs.data.inspection}
if(settings$generate.paper.figs) {

# Dimension of data after analysis
cat("Mixed-batch data:\n")
cat("  - Dimensions: p=", dim(test.RC)[1],", n=", dim(test.RC)[2], "\n", sep="")
cat("  - Positive controls:\n")
cat("    * Definition mean RC in interval: [", config$t.well, ", inf ]\n", sep="")
cat("    * Number of positive controls miRNAs:", length(pos.controls.test), "\n")
cat("  - Negative controls:\n")
cat("    * Definition mean RC in interval: [", config$t.zero, ",", config$t.poor, "]\n")
cat("    * Number of negative controls miRNAs:", length(neg.controls.test), "\n")

cat("\n\nSingle-batch data:\n")
cat("  - Dimensions: p=", dim(bench.RC)[1],", n=", dim(bench.RC)[2], "\n", sep="")
cat("  - Positive controls:\n")
cat("    * Definition mean RC in interval: [", config$t.well, ", inf ]\n", sep="")
cat("    * Number of positive controls miRNAs:", length(pos.controls.bench), "\n")
cat("  - Negative controls:\n")
cat("    * Definition mean RC in interval: [", config$t.zero, ",", config$t.poor, "]\n")
cat("    * Number of negative controls miRNAs:", length(neg.controls.bench), "\n")

}
```

## Read Count Distribution in the Data

### Per Data-Set

```{r generate.paper.figs.RC.distribution}
if(settings$generate.paper.figs) {
  par(mfrow=c(1,1))

  # Histogram plot test data set
  test.RC.log2 <- log2(rowMeans(test.RC)+1)
  df <- data.frame(log.expression=test.RC.log2)
  p.test.RC.hist <- ggplot(df, aes(x=log.expression)) + 
    geom_histogram(binwidth = .1, color="black", fill="black") +
    geom_vline(xintercept = log2(config$t.zero+1), color="#3051ab", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.poor+1), color="#3051ab", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.well+1), color="9e086c",  linetype="longdash") +
    ggtitle("Test data set") +
    theme_classic()
  print(p.test.RC.hist)
  rm(df)

  # Histogram plot benchmark data set
  bench.RC.log2 <- log2(rowMeans(bench.RC)+1)
  df <- data.frame(log.expression=bench.RC.log2)
  p.bench.RC.hist <- ggplot(df, aes(x=log.expression)) + 
    geom_histogram(binwidth = .1, color="black", fill="black") +
    geom_vline(xintercept = log2(config$t.zero+1), color="#3051ab", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.poor+1), color="#3051ab", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.well+1), color="#9e086c",  linetype="longdash") +
    ggtitle("Benchmark data set") +
    theme_classic()
  print(p.bench.RC.hist)
  rm(df)

  # Combined read count histogram for test and benchmark data
  df <- data.frame(bench=bench.RC.log2, test=test.RC.log2)
  p.RC.hist <- ggplot(df) + 
    theme_classic() +
    geom_histogram(aes(x=test, y=..count..,fill="#69b3a2"), binwidth = .1) +
    geom_histogram(aes(x=bench, y=-..count.., fill="#404080"), binwidth = .1) +
    geom_vline(xintercept = log2(config$t.zero+1), color="#5851b8", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.poor+1), color="#5851b8", linetype="twodash") +
    geom_vline(xintercept = log2(config$t.well+1), color="#E7298A",  linetype="longdash") +
    geom_segment(aes(x = log2(config$t.zero+1), y = 180, xend = log2(config$t.poor+1), yend = 180),
                 arrow=arrow(length=unit(.07, "in"), ends="both"),
                 color="#5851b8") +
    annotate(geom="label", x=(log2(config$t.zero+1)+log2(config$t.poor+1))/2.0, y=200,
             label="neg. contr.", color="#5851b8", size=4, fill = alpha(c("white"),0.85), label.size = NA) +
    geom_segment(aes(x = log2(config$t.well+1), y = 180, xend = 10, yend = 180),
                 arrow=arrow(length=unit(.07, "in"), ends="last"),
                 color="#E7298A") +
    annotate(geom="label", x=log2(config$t.well+1)+.5, y=200,
             label="pos. contr.", color="#E7298A", size=4, hjust=0, fill = alpha(c("white"),0.85), label.size = NA) +
    # ylim(c(-90,90)) +
    scale_fill_manual(name="Data set",values=c("#404080", "#69b3a2"), labels=c("single-batch", "mixed-batch"), guide = guide_legend(reverse=TRUE)) +
    # ggtitle("log2(Read count + 1)") +
    theme(legend.position=c(0.86,0.86), 
          # legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.9))) +
    scale_y_continuous(breaks=c(-200,-150,-100,-50,0,50,100,150,200), labels=c(200,150,100,50,0,50,100,150,200), limits=c(-200,200)) +
    xlab("Mean") +
    ylab("Frequency")
  print(p.RC.hist)
  rm(df)

    ## Export Plots
  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_RC_Distribution.pdf"), p.RC.hist, width = 5, height=4, device="pdf")
  }

}
```

### Group Means comparison between mixed- and single-batch data

```{r generate.paper.figs.RC.comparison}
if(settings$generate.paper.figs) {
  par(mfrow=c(1,1))

  ## log2(mean(RC))

  # Histogram plot test data set
  test.RC.log2 <- log2(rowMeans(test.RC)+1)
  bench.RC.log2 <- log2(rowMeans(bench.RC)+1)
  df <- data.frame(test=test.RC.log2, benchmark=bench.RC.log2)
  p.RC.comparison.1 <- ggplot(df,aes(x=test,y=benchmark)) + 
    geom_point(alpha=.25) + 
    xlab("mixed-batch") + 
    ylab("single-batch") + 
    ggtitle("log2(mean(read count) + 1)") +
      geom_abline(intercept = 0, slope = 1, color="black", linetype="longdash", size=.5) +
    theme(aspect.ratio = 1) +
    theme_classic()
  print(p.RC.comparison.1)
  rm(df,test.RC.log2, bench.RC.log2)

  # Histogram plot test data set (END only)
  test.RC.log2 <- log2(rowMeans(test.RC[, test.groups=="END"])+1)
  bench.RC.log2 <- log2(rowMeans(bench.RC[, bench.groups=="END"])+1)
  df <- data.frame(test=test.RC.log2, benchmark=bench.RC.log2)
  p.RC.comparison.END.1 <- ggplot(df,aes(x=test,y=benchmark)) + 
    geom_point(alpha=.25) + 
    xlab("mixed-batch (END)") + 
    ylab("single-batch (END)") + 
    ggtitle(" END  ---  log2(mean(read count) + 1)") +
      geom_abline(intercept = 0, slope = 1, color="black", linetype="longdash", size=.5) +
    theme(aspect.ratio = 1) +
    theme_classic()
  print(p.RC.comparison.END.1)
  rm(df,test.RC.log2, bench.RC.log2)

  # Histogram plot test data set (SER only)
  test.RC.log2 <- log2(rowMeans(test.RC[, test.groups=="SER"])+1)
  bench.RC.log2 <- log2(rowMeans(bench.RC[, bench.groups=="SER"])+1)
  df <- data.frame(test=test.RC.log2, benchmark=bench.RC.log2)
  p.RC.comparison.SER.1 <- ggplot(df,aes(x=test,y=benchmark)) + 
    geom_point(alpha=.25) + 
    xlab("mixed-batch (SER)") + 
    ylab("single-batch (SER)") + 
    ggtitle(" SER  ---  log2(mean(read count) + 1)") +
      geom_abline(intercept = 0, slope = 1, color="black", linetype="longdash", size=.5) +
    theme(aspect.ratio = 1) +
    theme_classic()
  print(p.RC.comparison.SER.1)
  rm(df,test.RC.log2, bench.RC.log2)

  ## log2(mean(RC))

  # Histogram plot test data set
  test.RC.log2 <- rowMeans(log2(test.RC+1))
  bench.RC.log2 <- rowMeans(log2(bench.RC+1))
  df <- data.frame(test=test.RC.log2, benchmark=bench.RC.log2)
  p.RC.comparison.2 <- ggplot(df,aes(x=test,y=benchmark)) + 
    geom_point(alpha=.25) + 
    xlab("mixed-batch") + 
    ylab("single-batch") + 
    ggtitle("mean(log2(read count + 1))") +
      geom_abline(intercept = 0, slope = 1, color="black", linetype="longdash", size=.5) +
    theme(aspect.ratio = 1) +
    theme_classic()
  print(p.RC.comparison.2)
  rm(df,test.RC.log2, bench.RC.log2)

  # Histogram plot test data set (END only)
  test.RC.log2 <- rowMeans(log2(test.RC[, test.groups=="END"]+1))
  bench.RC.log2 <- rowMeans(log2(bench.RC[, bench.groups=="END"]+1))
  df <- data.frame(test=test.RC.log2, benchmark=bench.RC.log2)
  p.RC.comparison.END.2 <- ggplot(df,aes(x=test,y=benchmark)) + 
    geom_point(alpha=.25) + 
    xlab("mixed-batch (END)") + 
    ylab("single-batch (END)") + 
    ggtitle(" END  ---  mean(log2(read count + 1))") +
      geom_abline(intercept = 0, slope = 1, color="black", linetype="longdash", size=.5) +
    theme(aspect.ratio = 1) +
    theme_classic()
  print(p.RC.comparison.END.2)
  rm(df,test.RC.log2, bench.RC.log2)

  # Histogram plot test data set (SER only)
  test.RC.log2 <- rowMeans(log2(test.RC[, test.groups=="SER"]+1))
  bench.RC.log2 <- rowMeans(log2(bench.RC[, bench.groups=="SER"]+1))
  df <- data.frame(test=test.RC.log2, benchmark=bench.RC.log2)
  p.RC.comparison.SER.2 <- ggplot(df,aes(x=test,y=benchmark)) + 
    geom_point(alpha=.25) + 
    xlab("mixed-batch (SER)") + 
    ylab("single-batch (SER)") + 
    ggtitle(" SER  ---  mean(log2(read count + 1))") +
      geom_abline(intercept = 0, slope = 1, color="black", linetype="longdash", size=.5) +
    theme(aspect.ratio = 1) +
    theme_classic()
  print(p.RC.comparison.SER.2)
  rm(df,test.RC.log2, bench.RC.log2)

    ## Export Plots
  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_RC_Comparison_all.pdf"), p.RC.comparison.2, width = 5, height=5, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_RC_Comparison_END.pdf"), p.RC.comparison.END.2, width = 5, height=5, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_RC_Comparison_SER.pdf"), p.RC.comparison.SER.2, width = 5, height=5, device="pdf")
  }

}
```

## Mean-Variance Plot

### Distribution property among technical replicates

```{r generate.paper.figs.mean.variance.plot1}
if(settings$generate.paper.figs) {
  # Function for mean-variance plot for a given data set
  mean.variance.plot <- function(data.RC, title, axis.limit=NULL) {
    df <- data.frame(data.mean=log10(rowMeans(data.RC) + 1),
                     data.var =log10(rowVars(data.RC) + 1))
    lin.fit <- lm(df$data.var ~ df$data.mean)$coefficients
    p <- ggplot(df,aes(x=data.mean,y=data.var)) + 
      geom_point(alpha=.25) + 
      xlab("log10(Mean)") + 
      ylab("log10(Variance)") + 
      geom_abline(intercept = lin.fit[1], slope = lin.fit[2], color="red", linetype="longdash", size=1) +
      geom_abline(intercept = 0, slope = 1, colour="blue") +
      annotate("text", alpha=1, x=4, y=0.5, hjust=0, vjust=0, size=3, colour="red",
               label=paste0("log10(Variance) = ", round(lin.fit[1],2), " + ", round(lin.fit[2],2), "*log10(Mean)")) + 
      xlim(0, ifelse(is.null(axis.limit), max(df), axis.limit)) +
      ylim(0, ifelse(is.null(axis.limit), max(df), axis.limit)) +
      ggtitle(title) +
      theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) +
      theme_classic()
    return(p)
  }

  ## Test data
  # END and SER
  p.meanvar.test <- mean.variance.plot(test.RC, "Mixed-batch Data", axis.limit=12.5)
  print(p.meanvar.test)
  # END (endometrioid)
  p.meanvar.test.END <- mean.variance.plot(test.RC[, test.groups == "END"], "Mixed-batch Data - END (endometrioid)", axis.limit=12.5)
  print(p.meanvar.test.END)
  # SER (serous)
  p.meanvar.test.SER <- mean.variance.plot(test.RC[, test.groups == "SER"], "Mixed-batch Data - SER (serous)", axis.limit=12.5)
  print(p.meanvar.test.SER)

  ## Benchmark data
  # END and SER
  p.meanvar.bench <- mean.variance.plot(bench.RC, "Single-batch Data", axis.limit=12.5)
  print(p.meanvar.bench)
  # END
  p.meanvar.bench.END <- mean.variance.plot(bench.RC[, bench.groups == "END"], "Single-batch Data - END (endometrioid)", axis.limit=12.5)
  print(p.meanvar.bench.END)
  # SER
  p.meanvar.bench.SER <- mean.variance.plot(bench.RC[, bench.groups == "SER"], "Single-batch Data - SER (serous)", axis.limit=12.5)
  print(p.meanvar.bench.SER)
}
```

### Marker-specific sd vs mean

```{r generate.paper.figs.mean.variance.plot2}
if(settings$generate.paper.figs) {

  ## Mixed-batch data
  # END and SER
  p.meanvar.test <- plot.mean.sd(test.RC, config$t.zero, config$t.poor, config$t.well)
  print(p.meanvar.test)
  # END (endometrioid)
  p.meanvar.test.END <- plot.mean.sd(test.RC[, test.groups == "END"], config$t.zero, config$t.poor, config$t.well, "Mixed-batch Data - END (endometrioid)")
  print(p.meanvar.test.END)
  # SER
  p.meanvar.test.SER <- plot.mean.sd(test.RC[, test.groups == "SER"], config$t.zero, config$t.poor, config$t.well, "Mixed-batch Data - SER (serous)")
  print(p.meanvar.test.SER)

  ## Single-Batch data
  # END and SER
  p.meanvar.bench <- plot.mean.sd(bench.RC, config$t.zero, config$t.poor, config$t.well, "Single-Batch Data")
  print(p.meanvar.bench)
  # END (endometrioid)
  p.meanvar.bench.END <- plot.mean.sd(bench.RC[, bench.groups == "END"], config$t.zero, config$t.poor, config$t.well, "Single-Batch Data - END (endometrioid)")
  print(p.meanvar.bench.END)
  # SER
  p.meanvar.bench.SER <- plot.mean.sd(bench.RC[, bench.groups == "SER"], config$t.zero, config$t.poor, config$t.well, "Single-Batch Data - SER (serous)")
  print(p.meanvar.bench.SER)

  ## Export Plots
  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_Test_MeanSD.pdf"), p.meanvar.test + labs(title = NULL), width = 5, height=4, device="pdf")
  }
}
```

## Correlation Plots

```{r generate.paper.figs.correlation}
if(settings$generate.paper.figs) {
  num.genes.plot.pos <- 60
  num.genes.plot.neg <- 60

  # Co-expression models
  bench.model <- DANA.bench$data.model
  test.model <- DANA.test$data.model
  test.norm.model <- DANA.test$norm.models$test.TMM

  # Common control miRNAs
  pos.controls.bench <- rownames(bench.model$pos$corr)
  pos.controls.test <- rownames(test.model$pos$corr)
  pos.controls.common <- intersect(pos.controls.test, pos.controls.bench)
  neg.controls.bench <- rownames(bench.model$neg$corr)
  neg.controls.test <- rownames(test.model$neg$corr)
  # neg.controls.common <- intersect(neg.controls.test, neg.controls.bench)
  # reduce the number of genes for the plot
  num.genes.plot.pos <- min(num.genes.plot.pos, length(pos.controls.common))
  pos.controls.plot <- pos.controls.common[1:num.genes.plot.pos]

  num.genes.plot.neg <- min(num.genes.plot.neg,
                            dim(test.model$neg$corr)[1],
                            dim(bench.model$neg$corr)[1])

  # Subsetted correlation matrices
  corr.bench.pos <- bench.model$pos$corr[pos.controls.plot, pos.controls.plot]
  corr.test.pos <- test.model$pos$corr[pos.controls.plot, pos.controls.plot]
  corr.test.TMM.pos <- test.norm.model$pos$corr[pos.controls.plot, pos.controls.plot]
  corr.bench.neg <- bench.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.neg <- test.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.TMM.neg <- test.norm.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]

  ### Generate plots
  # Positive controls
  p.corr.pos.bench <- ggplot.corr(corr.bench.pos, 
                                  clusters=clusters, 
                                  threshold=0, 
                                  title="Single-batch (un-normalized)",
                                  coords=coords)
  p.corr.pos.test <- ggplot.corr(corr.test.pos, 
                                 clusters=clusters, 
                                 threshold=0, 
                                 title="Mixed-batch (un-normalized)",
                                 coords=coords)
  p.corr.pos.test.TMM <- ggplot.corr(corr.test.TMM.pos, 
                                     clusters=clusters, 
                                     threshold=0, 
                                     title="Mixed-batch (TMM)",
                                     coords=coords)

  p.corr.pos <- arrangeGrob(p.corr.pos.bench + theme(legend.position="none"), 
                            p.corr.pos.test + theme(legend.position="none"),
                            get.legend(p.corr.pos.bench), 
                            widths = c(2,2,1))
  grid.arrange(p.corr.pos)  # plot

  p.corr.pos.three <- arrangeGrob(p.corr.pos.bench + theme(legend.position="none"), 
                                  p.corr.pos.test.TMM + theme(legend.position="none"),
                                  p.corr.pos.test + theme(legend.position="none"),
                                  get.legend(p.corr.pos.bench + theme(legend.position = "bottom")),
                                  layout_matrix=rbind(c(1,2,3), c(4,4,4)),
                                  heights = c(5,1))
  grid.arrange(p.corr.pos.three)  # plot

  # Negative controls
  p.corr.neg.bench <- ggplot.corr(corr.bench.neg, 
                                  clusters=clusters, 
                                  threshold=0, 
                                  title="Single-batch (un-normalized)")
  p.corr.neg.test <- ggplot.corr(corr.test.neg, 
                                 clusters=clusters, 
                                 threshold=0, 
                                 title="Mixed-batch (un-normalized)")
  p.corr.neg.test.TMM <- ggplot.corr(corr.test.TMM.neg, 
                                     clusters=clusters, 
                                     threshold=0, 
                                     title="Mixed-batch (TMM)")

  p.corr.neg <- arrangeGrob(p.corr.neg.bench + theme(legend.position="none"), 
                            p.corr.neg.test + theme(legend.position="none"),
                            get.legend(p.corr.neg.bench),
                            widths = c(2,2,1))
  grid.arrange(p.corr.neg)  # plot

  p.corr.neg.three <- arrangeGrob(p.corr.neg.bench + theme(legend.position="none"), 
                                  p.corr.neg.test.TMM + theme(legend.position="none"),
                                  p.corr.neg.test + theme(legend.position="none"),
                                  get.legend(p.corr.neg.bench + theme(legend.position = "bottom")),
                                  layout_matrix=rbind(c(1,1,2,2,3,3), c(4,4,4,4,4,4)),
                                  heights = c(5,1))
  grid.arrange(p.corr.neg.three)  # plot

  ## Export Plots
  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_CorrPos.pdf"), p.corr.pos, width = 8, height=3.5, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_CorrNeg.pdf"), p.corr.neg, width = 8, height=3.5, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_CorrPos_TMM.pdf"), p.corr.pos.three, width = 8, height=3.5, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_CorrNeg_TMM.pdf"), p.corr.neg.three, width = 8, height=3.5, device="pdf")
  }
}
```

### All normalization methods

```{r generate.paper.figs.correlation.all}
if(settings$generate.paper.figs) { 
  p.corr.test <- ggplot.corr(DANA.test$data.model$pos$corr,
                             clusters=clusters,
                             title="un-normalized")
  p.corr.bench <- ggplot.corr(DANA.bench$data.model$pos$corr,
                              clusters=clusters,
                              title="benchmark data (un-normalized)")
  p.corr.TMM <- ggplot.corr(DANA.test$norm.models$test.TMM$pos$corr,
                            clusters=clusters,
                            title="TMM")
  p.corr.DESeq <- ggplot.corr(DANA.test$norm.models$test.DESeq$pos$corr,
                            clusters=clusters,
                            title="DESeq")
  p.corr.PoissonSeq <- ggplot.corr(DANA.test$norm.models$test.PoissonSeq$pos$corr,
                            clusters=clusters,
                            title="PoissonSeq")
  p.corr.TC <- ggplot.corr(DANA.test$norm.models$test.TC$pos$corr,
                            clusters=clusters,
                            title="TC")
  p.corr.Med <- ggplot.corr(DANA.test$norm.models$test.Med$pos$corr,
                            clusters=clusters,
                            title="Med")
  p.corr.RUVg <- ggplot.corr(DANA.test$norm.models$test.RUVg$pos$corr,
                            clusters=clusters,
                            title="RUVg")
  p.corr.RUVs <- ggplot.corr(DANA.test$norm.models$test.RUVs$pos$corr,
                            clusters=clusters,
                            title="RUVs")
  p.corr.RUVr <- ggplot.corr(DANA.test$norm.models$test.RUVr$pos$corr,
                            clusters=clusters,
                            title="RUVr")
  p.corr.UQ <- ggplot.corr(DANA.test$norm.models$test.UQ$pos$corr,
                            clusters=clusters,
                            title="UQ")
  p.corr.QN <- ggplot.corr(DANA.test$norm.models$test.QN$pos$corr,
                            clusters=clusters,
                            title="QN")

  # Arrange plots
  p.corr.test.all <-
    plot_grid(p.corr.test + theme(legend.position="none"),
              p.corr.TMM + theme(legend.position="none"),
              p.corr.DESeq + theme(legend.position="none"),
              p.corr.PoissonSeq + theme(legend.position="none"),
              p.corr.TC + theme(legend.position="none"),
              p.corr.Med + theme(legend.position="none"),
              p.corr.RUVg + theme(legend.position="none"),
              p.corr.RUVs + theme(legend.position="none"),
              p.corr.RUVr + theme(legend.position="none"),
              p.corr.UQ + theme(legend.position="none"),
              p.corr.QN + theme(legend.position="none"),
              get.legend(p.corr.test + theme(legend.position = "bottom")),
              ncol=3)
  plot(p.corr.test.all)

  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_Test_Corr_All.pdf"), p.corr.test.all, width=9, height=12, device="pdf")
  }
}
```

## Correlation Frequency Plots

### Subset for paper

```{r generate.paper.figs.corr.density}
if(settings$generate.paper.figs) {
  # Co-expression models
  bench.model <- DANA.bench$data.model
  bench.norm.model <- DANA.bench$norm.models$benchmark.TMM
  test.model <- DANA.test$data.model
  test.norm.model1 <- DANA.test$norm.models$test.TMM
  test.norm.model2 <- DANA.test$norm.models$test.TC

  # Set number of genes for negative correlation to minimum
  num.genes.plot.neg <- min(dim(test.model$neg$corr)[1],
                            dim(bench.model$neg$corr)[1],
                            dim(test.norm.model1$neg$corr)[1],
                            dim(test.norm.model2$neg$corr)[1])

  # Subsetted correlation matrices
  corr.bench.neg <- bench.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.bench.norm1.neg <- bench.norm.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.neg <- test.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.norm1.neg <- test.norm.model1$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.norm2.neg <- test.norm.model2$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]

  ## extract non-zero correlations
  nnz.corr.test.neg <- corr.test.neg[upper.tri(corr.test.neg)]
  nnz.corr.test.neg <- nnz.corr.test.neg[nnz.corr.test.neg != 0]
  nnz.corr.test.norm1.neg <- corr.test.norm1.neg[upper.tri(corr.test.norm1.neg)]
  nnz.corr.test.norm1.neg <- nnz.corr.test.norm1.neg[nnz.corr.test.norm1.neg != 0]
  nnz.corr.test.norm2.neg <- corr.test.norm2.neg[upper.tri(corr.test.norm2.neg)]
  nnz.corr.test.norm2.neg <- nnz.corr.test.norm2.neg[nnz.corr.test.norm2.neg != 0]
  nnz.corr.bench.neg <- corr.bench.neg[upper.tri(corr.bench.neg)]
  nnz.corr.bench.neg <- nnz.corr.bench.neg[nnz.corr.bench.neg != 0]
  nnz.corr.bench.norm1.neg <- corr.bench.norm1.neg[upper.tri(corr.bench.norm1.neg)]
  nnz.corr.bench.norm1.neg <- nnz.corr.bench.norm1.neg[nnz.corr.bench.norm1.neg != 0]

  # Keep positive correlations
  nnz.corr.test.neg <- nnz.corr.test.neg[nnz.corr.test.neg > 0]
  nnz.corr.test.norm1.neg <- nnz.corr.test.norm1.neg[nnz.corr.test.norm1.neg > 0]
  nnz.corr.test.norm2.neg <- nnz.corr.test.norm2.neg[nnz.corr.test.norm2.neg > 0]
  nnz.corr.bench.neg <- nnz.corr.bench.neg[nnz.corr.bench.neg > 0]
  nnz.corr.bench.norm1.neg <- nnz.corr.bench.norm1.neg[nnz.corr.bench.norm1.neg > 0]

  ## direct comparison of negative controls benchmark vs test vs test.norm1
  neg.corr <- data.frame(
    control=factor(c(rep("Single-batch", length(nnz.corr.bench.neg)), 
                     rep("Mixed-batch", length(nnz.corr.test.neg)),
                     rep("Mixed-batch (TMM)", length(nnz.corr.test.norm1.neg)))),
    corr=abs(c(nnz.corr.bench.neg, nnz.corr.test.neg, nnz.corr.test.norm1.neg))
    )

  p.density.neg.nnz.freq <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.05, alpha=.8, size=.9) +
    theme_classic() +
    xlim(0, 1) +
    scale_linetype_manual(values=c("solid", "solid", "twodash"))+
    scale_color_manual(values=c('#1B9E77','#D95F02','#D95F02')) + # First 2 colors from rcolorbrewer set "Dark2"
    theme(legend.position=c(0.9,0.9), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    ylab("Frequency") +
    xlab("Marginal correlation") +
    ggtitle("Positive correlation among negative controls")
  print(p.density.neg.nnz.freq)

  ## direct comparison of negative controls Single-batch vs test vs test.norm1 vs test.norm2
  neg.corr <- data.frame(
    control=factor(c(rep("Single-batch", length(nnz.corr.bench.neg)), 
                     rep("Mixed-batch", length(nnz.corr.test.neg)),
                     rep("Mixed-batch (TMM)", length(nnz.corr.test.norm1.neg)),
                     rep("Mixed-batch (TC)", length(nnz.corr.test.norm2.neg)))),
    corr=abs(c(nnz.corr.bench.neg, nnz.corr.test.neg, nnz.corr.test.norm1.neg, nnz.corr.test.norm2.neg))
    )
  # ttt <- data.frame(prop.table(table(method=neg.corr$control, corr=neg.corr$corr), 1))

  p.density.neg.nnz.freq <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.05, alpha=.8, size=.9) +
    theme_classic() +
    xlim(0, 1) +
    # scale_linetype_manual(values=c("twodash", "solid", "solid", "solid"))+
    scale_color_manual(values=c('#e7298a','#1B9E77','#D95F02', '#7570B3')) + # First colors from rcolorbrewer set "Dark2"
    theme(legend.position=c(0.9,0.86), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    ylab("Frequency") +
    xlab("Marginal correlation")#+ ggtitle("Positive correlation among negative controls")
  print(p.density.neg.nnz.freq)

  # Additional Plot with normalized Single-batch
  neg.corr <- data.frame(
    control=factor(c(rep("Single-batch", length(nnz.corr.bench.neg)), 
                     rep("Single-batch (TMM)", length(nnz.corr.bench.norm1.neg)),
                     rep("Mixed-batch", length(nnz.corr.test.neg)),
                     rep("Mixed-batch (TMM)", length(nnz.corr.test.norm1.neg)))),
    corr=abs(c(nnz.corr.bench.neg, nnz.corr.bench.norm1.neg, nnz.corr.test.neg, nnz.corr.test.norm1.neg))
    )
  p.density.neg.nnz.freq.four <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.05, alpha=.8, size=1) +
    theme_classic() +
    xlim(0, 1) +
    theme(legend.position=c(0.9,0.9), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    ylab("Frequency") +
    xlab("Marginal correlation") +
    scale_linetype_manual(values=c("solid", "twodash","solid", "twodash"))+
    scale_color_manual(values=c('#1B9E77','#1B9E77','#D95F02','#D95F02')) + # First 2 colors from rcolorbrewer set "Dark2"
    ggtitle("Positive correlation among negative controls")
  print(p.density.neg.nnz.freq.four)

  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_CorrDensityNeg_Freq.pdf"), p.density.neg.nnz.freq, width = 5, height=4, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_CorrDensityNeg_Freq_WithBench.pdf"), p.density.neg.nnz.freq.four, width = 5, height=4, device="pdf")
  }
}
```

### All methods

```{r generate.paper.figs.corr.density.supplements}
if(settings$generate.paper.figs) {
  # Co-expression models
  bench.model <- DANA.bench$data.model
  test.model <- DANA.test$data.model
  test.norm.model1 <- DANA.test$norm.models$test.TMM
  test.norm.model2 <- DANA.test$norm.models$test.RUVs

  # Set number of genes for negative correlation to minimum
  num.genes.plot.neg <- min(dim(test.model$neg$corr)[1],
                            dim(bench.model$neg$corr)[1],
                            sapply(DANA.test$norm.models, function(x) dim(x$neg$corr)[1]))

  # Subsetted correlation matrices
  corr.bench.neg <- bench.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.neg <- test.model$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg]
  corr.test.norm.neg <- lapply(DANA.test$norm.models, function(x) x$neg$corr[1:num.genes.plot.neg, 1:num.genes.plot.neg])

  ## Strictly positive correlations
  nnz.corr.bench.neg <- corr.bench.neg[upper.tri(corr.bench.neg)]
  nnz.corr.bench.neg <- nnz.corr.bench.neg[nnz.corr.bench.neg > 0]
  nnz.corr.bench.norm1.neg <- corr.bench.norm1.neg[upper.tri(corr.bench.norm1.neg)]
  nnz.corr.bench.norm1.neg <- nnz.corr.bench.norm1.neg[nnz.corr.bench.norm1.neg > 0]
  nnz.corr.test.neg <- corr.test.neg[upper.tri(corr.test.neg)]
  nnz.corr.test.neg <- nnz.corr.test.neg[nnz.corr.test.neg > 0]
  nnz.corr.test.norm.neg <- lapply(corr.test.norm.neg, function(x) x[upper.tri(x)][x[upper.tri(x)] > 0])
  names(nnz.corr.test.norm.neg) <- substr(names(nnz.corr.test.norm.neg), 6, 20)

  control <- c(rep("Single-batch", length(nnz.corr.bench.neg)), 
                     rep("Mixed-batch", length(nnz.corr.test.neg)))
  for(i in 1:length(nnz.corr.test.norm.neg)) {
    control <- c(control, rep(paste0("Mixed-batch (", names(nnz.corr.test.norm.neg)[i],")"), length(nnz.corr.test.norm.neg[[i]])))
  }
  ## direct comparison of negative controls Single-batch vs test vs test.norm1
  neg.corr <- data.frame(
    control=factor(control),
    corr=abs(c(nnz.corr.bench.neg, nnz.corr.test.neg, unname(unlist(nnz.corr.test.norm.neg))))
    )

  p.corr.frequency.neg.controls.all <- ggplot(neg.corr, aes(x=corr, colour=control, linetype=control)) +
    geom_freqpoly(binwidth=.05, alpha=.9, size=.9) +
    theme_classic() +
    xlim(0, 1) +
    scale_color_manual(values=brewer.pal(n=12, name="Paired")) + # First 2 colors from rcolorbrewer set "Dark2"
    theme(legend.position=c(0.85,0.6), 
          legend.title=element_blank(),
          legend.background=element_rect(fill=alpha('white', 0.5))) +
    ylab("Frequency") +
    xlab("Marginal correlation") + # ggtitle("Positive correlation among negative controls") + 
    theme(legend.key.width = unit(2,"cm"))
  print(p.corr.frequency.neg.controls.all)

  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_CorrDensityNeg_Freq_All.pdf"), p.corr.frequency.neg.controls.all, width = 8, height=5, device="pdf")
  }
}
```

## Correlation Scatter Plots

### Subset for paper

```{r generate.paper.figs.corr.scatter}
if(settings$generate.paper.figs) {
  ## Generate correlation scatter plots for positive controls for TMM and UQ
  par(mfrow=c(1,1))

  # un-normalized vs TMM
  p.scatter.test.TMM <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.TMM$pos$corr,
    clusters=clusters.test,
    title="Mixed-batch (TMM)",
    xlab="un-normalized", 
    ylab="TMM", 
    threshold=0)
  print(p.scatter.test.TMM)
  # un-normalized vs RUVs
  p.scatter.test.RUVs <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.RUVs$pos$corr,
    clusters=clusters.test,
    title="Mixed-batch (RUVs)",
    xlab="un-normalized", 
    ylab="RUVs", 
    threshold=0)
  print(p.scatter.test.RUVs)

  # Save plots
  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_ScatterTMM.pdf"), p.scatter.test.TMM, width=4, height=4, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_ScatterRUVs.pdf"), p.scatter.test.RUVs, width=4, height=4, device="pdf")
    # ggsave(paste0(settings$paper.fig.path, "UCEC_CorrNeg.eps"), p.corr.neg, width = 8, height=3.5, device="eps")
  }
}
```

### All normalization methods for supplementary materials

```{r generate.paper.figs.corr.scatter.all}
if(settings$generate.paper.figs) {
  par(mfrow=c(1,1))

  # un-normalized vs TMM
  p.scatter.test.TMM <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.TMM$pos$corr,
    clusters=clusters.test, title="TMM", xlab="un-normalized", ylab="TMM", 
    ccc=round(DANA.test$metrics["test.TMM", "cc"],3))
  # un-normalized vs DESeq
  p.scatter.test.DESeq <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.DESeq$pos$corr,
    clusters=clusters.test, title="DESeq", xlab="un-normalized", ylab="DESeq", 
    ccc=round(DANA.test$metrics["test.DESeq", "cc"],3))
  # un-normalized vs PoissonSeq
  p.scatter.test.PoissonSeq <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.PoissonSeq$pos$corr,
    clusters=clusters.test, title="PoissonSeq", xlab="un-normalized", ylab="PoissonSeq", 
    ccc=round(DANA.test$metrics["test.PoissonSeq", "cc"],3))
  # un-normalized vs TC
  p.scatter.test.TC <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.TC$pos$corr,
    clusters=clusters.test, title="TC", xlab="un-normalized", ylab="TC", 
    ccc=round(DANA.test$metrics["test.TC", "cc"],3))
  # un-normalized vs Med
  p.scatter.test.Med <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.Med$pos$corr,
    clusters=clusters.test, title="Med", xlab="un-normalized", ylab="Med", 
    ccc=round(DANA.test$metrics["test.Med", "cc"],3))
  # un-normalized vs RUVg
  p.scatter.test.RUVg <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.RUVg$pos$corr,
    clusters=clusters.test, title="RUVg", xlab="un-normalized", ylab="RUVg", 
    ccc=round(DANA.test$metrics["test.RUVg", "cc"],3))
  # un-normalized vs RUVs
  p.scatter.test.RUVs <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.RUVs$pos$corr,
    clusters=clusters.test, title="RUVs", xlab="un-normalized", ylab="RUVs", 
    ccc=round(DANA.test$metrics["test.RUVs", "cc"],3))
  # un-normalized vs RUVr
  p.scatter.test.RUVr <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.RUVr$pos$corr,
    clusters=clusters.test, title="RUVr", xlab="un-normalized", ylab="RUVr", 
    ccc=round(DANA.test$metrics["test.RUVr", "cc"],3))
  # un-normalized vs UQ
  p.scatter.test.UQ <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.UQ$pos$corr,
    clusters=clusters.test, title="UQ", xlab="un-normalized", ylab="UQ", 
    ccc=round(DANA.test$metrics["test.UQ", "cc"],3))
  # un-normalized vs QN
  p.scatter.test.QN <- plot.corr.scatter(
    corr1=DANA.test$data.model$pos$corr, 
    corr2=DANA.test$norm.models$test.QN$pos$corr,
    clusters=clusters.test, title="QN", xlab="un-normalized", ylab="QN", 
    ccc=round(DANA.test$metrics["test.QN", "cc"],3))

  p.scatter.test.all <- plot_grid(p.scatter.test.TMM,
                                  p.scatter.test.DESeq,
                                  p.scatter.test.PoissonSeq,
                                  p.scatter.test.TC,
                                  p.scatter.test.Med,
                                  p.scatter.test.RUVg,
                                  p.scatter.test.RUVs,
                                  p.scatter.test.RUVr,
                                  p.scatter.test.UQ,
                                  p.scatter.test.QN,
                                  ncol = 3,
                                  align="hv")
  plot(p.scatter.test.all)

  # Save plots
  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_Scatter_All.pdf"), p.scatter.test.all, width=9, height=12, device="pdf")
  }
}
```

## DANA Result Metrics

```{r generate.paper.figs.result.metrics}
if(settings$generate.paper.figs) { 
  options(scipen=4) # force non-scientific notation of x axis

  test.DANA.metrics <- data.frame(
    method=substr(rownames(DANA.test$metrics), 6, 20),
    cc=DANA.test$metrics[, "cc"],
    mcr=DANA.test$metrics[, "mcr"]
  )

  p.metrics.test <- ggplot(test.DANA.metrics,aes(x=mcr,y=cc, label=method)) + 
    geom_point(alpha=.5) +
    theme_classic() + 
    xlab(TeX("$mcr^-$; Relative reduction of handling effects")) + 
    ylab(TeX("$cc^+$; Biological signal preservation")) + 
    geom_text_repel(aes(label = method), size=3) +
    # xlim(c(min(c(0, test.mposc.reduction)),max(test.mposc.reduction))) +
    scale_x_continuous(labels = scales::percent, limits=c(0,1.05), 
                       breaks = scales::pretty_breaks(n = 5)) +
    # ylim(c(0,1)) 
    scale_y_continuous(labels = scales::percent, limits = c(0,1))
  print(p.metrics.test)

  print("DANA Statistics for mixed-batch data")
  test.DANA.metrics$cc <- round(test.DANA.metrics$cc,3)
  test.DANA.metrics$mcr <- round(test.DANA.metrics$mcr,3)
  print(test.DANA.metrics)

  bench.DANA.metrics <- data.frame(
    method=substr(rownames(DANA.bench$metrics), 11, 30),
    cc=DANA.bench$metrics[, "cc"],
    mcr=DANA.bench$metrics[, "mcr"]
  )

  p.metrics.bench <- ggplot(bench.DANA.metrics, aes(x=mcr, y=cc, label=method)) + 
    geom_point(alpha=.5) +
    theme_classic() + 
    xlab(TeX("$mcr^-$; Relative reduction of handling effects")) + 
    ylab(TeX("$cc^+$; Biological signal preservation")) + 
    geom_text_repel(aes(label = method), size=3) +
    # xlim(c(min(c(0,bench.mposc.reduction)),max(bench.mposc.reduction))) +
    scale_x_continuous(labels = scales::percent, limits=c(0, 1.05),
                       breaks = scales::pretty_breaks(n = 5)) +
    # ylim(c(0,1))
    scale_y_continuous(labels = scales::percent, limits = c(0, 1)) 
  print(p.metrics.bench)

  print("DANA Statistics for single-batch data")
  bench.DANA.metrics$cc <- round(bench.DANA.metrics$cc,3)
  bench.DANA.metrics$mcr <- round(bench.DANA.metrics$mcr,3)
  print(bench.DANA.metrics)

  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_MetricsTest.pdf"), p.metrics.test, width=4.5, height=3.5, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_MetricsBench.pdf"), p.metrics.bench, width=4.5, height=3.5, device="pdf")
  }

}
```

## DEA Results

```{r generate.paper.figs.DEA}
if(settings$generate.paper.figs && settings$perform.DEA) { 
  # FNR-FDR Plot
  test.DEA.res <- data.frame(method=substr(names(test.norm.DEA.stats), 6, 20),
                             FDR=-sapply(test.norm.DEA.stats, function(x) x$FDR)+1,
                             FNR=-sapply(test.norm.DEA.stats, function(x) x$FNR)+1)
  p.test.DEA.FNR.FDR <- ggplot(test.DEA.res, aes(x=FNR, y=FDR, label=method)) +
      theme_classic() + 
      geom_point(alpha=1) + 
      xlab("Sensitivity") + # True positive rate(1-FNR)
      ylab("Precision") + # Positive predictive value (1-FDR)
      geom_text_repel(aes(label = method), size=3) +
      scale_x_continuous(labels = scales::percent, limits=c(0,1),
                         breaks = scales::pretty_breaks(n = 4)) +
      scale_y_continuous(labels = scales::percent, limits = c(0,1))
  print(p.test.DEA.FNR.FDR)

  # CAT Plot
  d <- append(list("test.No Norm"=test.DEA), test.norm.DEA)
  names(d) <- substr(names(d), 6, 20)
  p.test.DEA.CAT.1 <- plot.CAT(DEA=d[1:7], 
                             truth.DEA=bench.DEA, 
                             title="Significance ranks of miRNAs",
                             maxrank=100)
  print(p.test.DEA.CAT.1)
  p.test.DEA.CAT.2 <- plot.CAT(DEA=d[c(1,8:11)], 
                             truth.DEA=bench.DEA, 
                             title="Significance ranks of miRNAs",
                             maxrank=100)
  print(p.test.DEA.CAT.2)
  rm(d)

  if(settings$export.figures) {
    ggsave(paste0(settings$paper.fig.path, "UCEC_DEA_FNR_FDR.pdf"), p.test.DEA.FNR.FDR, width=4.5, height=3.5, device="pdf")
    ggsave(paste0(settings$paper.fig.path, "UCEC_DEA_CAT.pdf"), p.test.DEA.CAT, width=4.5, height=3.5, device="pdf")
  }
}
```

# Package Information

```{r session.info}
sessionInfo()
```